---
title: "EMS 206 Lab 5"
author: "Paul Rikhter"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives:
- Getting counts for different groups
- Use the {lubridate} package to parse dates
- Find confidence intervals and do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}

library(tidyverse)
library(janitor)
library(here)

```
```{r}
# Import lobster data

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()
```

Use the uncount function to expand a frequency variable
```{r}
lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```

Let's look at this a bunch of ways:
```{r}

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2)

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site))

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
   facet_wrap(~site, scales = "free")

```


Use 'luberdate' package to parse dates and times:
Will add a new column with mutate that contains the date as a date, not a character

```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Parse 'Date' column to get the different peices (month, year) separated
```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Count lobsters by different groupings
Let's say we want to count the number of lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_site <- lobster_parse_date %>% 
  count(site)

# group_by + summarise + n
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarise(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )
```

Confidence intervals and t-tests

Use the t.test function to find confidence intervals and perform t-tests

```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

my_vector <- c(3, 5, 9, 15, 24)

t.test(ivee_lobsters)

```

Two-sample t-test:
Is there a significant different in mean lobster lengths between Naples Reef and Mohawk Reef?

```{r}

naples_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohawk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_naple_ttest <- t.test(naples_sample, mohawk_sample)

```

Mean lobster size differed significantly between mohawk and naples reefs (t(`r mohk_naple_ttest$parameter`) = 19.849)

Maybe an easier way...
```{r}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL","MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)
```

Make a geom_time heatmap

```{r}
ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))


```


